Contextual Identity: Towards Usable Privacy

### 2013-03-04 paper.tex
Clean up punctuation, delete unnecessary words

\begin{abstract}
We examine \textit{contextual identity}, a notion that individuals reveal
different aspects of themselves depending on context. At any given time, they
may act as a friend, relative, spouse, co-worker, aquaintance, or stranger. We
analyze contextual identity from the perspective of user choice and control,
survey contextual identity violations, and propose new resesarch directions to
empower users to have better control over their contextual identities.
\end{abstract}

@@

The desire for privacy balances the desire for spontaneous, positive human
interaction, which often requires sharing personal information. The act of
sharing alone does not negate expectation of privacy~\cite{boyd2}.
%The desire for spontaneous, positive human interaction often requires sharing
%personal information. Sharing information can lead to loss of privacy, though
%it does not negate expectation of privacy~\cite{boyd2}. 
Helen Nissenbaum has long argued that privacy violations come not from the
simple sharing of personal information, but rather from sharing that
information in a way that violates social norms, or in the wrong
context~\cite{nissenbaum}.

Lesbian, gay, bisexual, and transgendered youth are at risk for contextual
identity violations.  Access to support is crucial for LGTB-identified youth,
who are at higher risk for bullying and suicide than non-LGTB youth~\cite{hrc}.
Much of that support is only accessible through the Internet in socially
conservative areas. In 2011 Bobbi Duncan and Taylor McCormick, two
students at the University of Texas at Austin, were inadvertantly outed when
their choir director added them to a Facebook group for Queer
Chorus~\cite{fowler}. Even though both students had blocked their parents from
seeing any personal posts revealing their sexuality, because Queer Chorus is a
public group Facebook published notifications to Duncan's and McCormick's
friends, including their parents. Duncan later attempted suicide~\cite{duncan}.

@@

Barth, Datta, Mitchell, and Nissenbaum provide a formal model for contextual
integrity of user data~\cite{barth}. The contextual integrity model focuses on
the flow of personal information in terms of knowledge states held by agents
acting in given roles, and works well for modeling privacy legislation such as
HIPPA and COPPA. Two limitations of this work are that it does not
model information where the subject involves more than one person, and it
only takes the type of information into account rather than the content or tone
of the information. For example, ``Bob told Alice that Charlie cooked a
delicious dinner'' is out of scope of the model.  In the context of social
networks, information is rarely about a single person, and the tone of the
message is as important to user effect as the contents.

@@

\label{sec:examples}
\subsection{Redistributing information out of context}
Re-distributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals and more
to the user's activity stream.  Because of poor opt-out and lack of visibility
into what was being published, user outcry was
immediate~\cite{mccarthy,nakashima}.  In December 2007, Google Reader exposed
RSS feeds of user-marked news stories to the user's Google Talk contacts, which
includes everyone with whom a user has chatted whether they be co-workers,
supervisors, or friends. Although this feed was always public, prior to this
launch it was not discoverable, leading to a bad experience for many
users~\cite{helft}. In February 2010, Google Buzz launched, a product that
exposed the user's most frequent Gmail and Google Talk contacts and their
publicly available (through previously less discoverable) news and photos.  The
combination of exposing contacts (which could include abusive ex-husbands,
co-workers, and friends), as well as aggressively linking photos and news
items, resulted in a huge backlash~\cite{fugitivus,buzz}. In September 2012,
Facebook imported old wall posts from 2008 into the new Timeline interface.
Although wall posts were always visible from profile pages, the new Timeline
interface brought old wall posts, which users used to treat as private messages
before the advent of ``Like'' and comment buttons, to the attention of an
audience that the user never anticipated~\cite{timeline}.

\subsection{Policy errors}
\label{sec:policies}
Some service providers have policies which preclude isolating multiple
contextual identities. For example, Facebook and Google have a ``Real Names''
policy, which requires users to register for accounts with their legal
name\footnote{Facebook's policy: \url{http://www.facebook.com/help/?page=258984010787183}, Google's policy: \url{http://support.google.com/plus/bin/answer.py?hl=en&answer=1228271}}.
%\cite{fb_names,google_names}.  
These policies ignore that
community-building happens in many different contexts, that users have many
legitimate reasons for presenting different identities in different contexts,
and that users don't necessarily want those identities to be linked. For
example, disallowing avatar handles as a primary identifier makes building a
gaming community difficult. % Linking a gaming identity to a professional
%identity has already caused problems for 2012 Maine Senate candidate and World
%of Warcraft gamer Colleen Lachowicz~\cite{maine}.
It is impossible to isolate multiple contextual identities in these networks
without violating the terms of service.

@@

Other federated login systems have support for multiple identifiers, but this
feature can be poorly implemented or difficult to discover, leading to low use.
Some login platforms and protocols, such as OAuth and BrowserID, do not suffer
from this policy or design error~\cite{browserid,oauth}, giving a user better
control over which contextual identity to assert.

\subsection{User interface errors}
It is all too easy for users to broadcast information to
an unintended audience. This phenomenon is so common on Twitter that it has its
own name, ``DM fail'', or Direct Message fail: when the user posts a public
message instead of a private, direct message. Representative Anthony Weiner was
a victim of this mistake when he inadvertently published compromising pictures
of himself~\cite{weiner}. Considering that this mistake requires
mistyping a single character (\texttt{@} instead of \texttt{d}), it's no
surprise that DM failures are exceedingly common.

@@

\subsection{How do users think about identity?}
We hypothesize many users are not consciously aware of having multiple
contextual identities. In order to develop tools to help users, we need to
understand users' mental models of identity and how information is shared on
the Internet. For computer scientists, authentication is tied to identity, and
even in the absence of authentication, using the same device over time
implicitly creates an identity through tracking techniques and local
information. However, these points are far from obvious to a typical user,
especially for users who don't distinguish the web application from the
browser, from the operating system, or from the device\footnote{What is a browser? \url{https://www.youtube.com/watch?v=o4MwTvtyrUQ}} \footnote{\url{http://input.mozilla.org/en-US/?q=facebook&product=firefox}}.

The computer science community must free itself from software-specific
notions of identity if we are to help users to whom the very concept of
identity is an enigma. To start understanding our users' mental models,
one might start with the questions:

@@

Pew Internet is a good source for phone survey data on privacy and social
media and danah boyd's work on youth and social media is excellent, but there
is otherwise a dearth of research in these areas~\cite{boyd,pew1,pew2,pew3}.

@@

Using multiple browsers is a useful technique for managing multiple accounts.
Some services support multiple accounts but not multiple login (e.g.,
Twitter). For services that do (e.g., many Google services), the user interface
may be so confusing that it leads to errors, so the best recommendation may be
to use multiple browsers~\cite{yegge}.\footnote{Google recommends multiple
browsers as an alternative to multiple login:
\url{https://support.google.com/accounts/bin/answer.py?hl=en&answer=179235}}.
Additionally, for users who want to separate work and
personal browsing, using multiple browsers is the easiest solution.

@@

Using multiple browsers forces separation between cookies, history, built-in
password managers, and other local storage. However, several browsers support
separation of profile data natively. Firefox 
%\url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}.}
and Chrome support multiple profiles, though neither of these is very
discoverable or easy to use.\footnote{Chrome: \url{https://support.google.com/chrome/bin/answer.py?hl=en&answer=2364824}, Firefox:
\url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}}
The threat model does not include users with malicious intent in
the same household.

\paragraph{Private browsing mode}
Private browsing mode exists in all major browsers, but has no standard
behavior~\cite{ABBJ10}.
Interaction with extensions, treatment of cookies, history, and
bookmarks upon entering and exiting are different across browsers.
\begin{comment}
For example, entering private browsing in Firefox does not clear
cookies, history, or other session data; session data is cleared on exiting
private browsing mode. In contrast, incognito mode in Chrome makes previous
session state unavailable and clears new session state on exit; this mode
may be preferable as a ``Guest mode'' where the device owner does not want the
guest to either view or pollute existing state.
\end{comment}
Similar to multiple profiles, private browsing mode is not secure against local
attacks.

\paragraph{Cookie blockers}
Disconnect and ShareMeNot are browser extensions to disallow interacting with
service providers like Google and Facebook unless users explicitly choose to
interact with that site~\cite{disconnect,franzi}.  Collusion is a tool for
visualizing cookies~\cite{collusion}, in particular third-party cookies set by
tracking sites which are typically ad networks.  Once the user knows which
sites are tracking them, they can choose to block cookies from those sites.

@@

Building reputation online is a long and arduous process; a user who has built
up high credibility in one social network may want to transfer that credibility
when using a new service by using the same identifier or somehow proving in the
new service that they control other existing, credible identities.  The rise of
verified account mechanisms in Twitter, Google Plus and Facebook indicates that
this is a real problem.

@@

All five major browsers (Firefox, Chrome, Safari, Internet Explorer, and Opera)
offer synchronizing some set of browser data across devices, with varying
degrees of success. Multiple cross-browser applications also exist to
synchronize bookmarks and passwords~\footnote{Bookmark sync: \url{xmarks.com}, password sync: \url{lastpass.com}}.

@@

Given the large number of social network users and that 30-40\% of spoken
communication is devoted to informing others about ourselves, many users are
bound to share personal information online that they regret~\cite{tamir}.

@@

\begin{comment}
\paragraph{Automatically inferring identities}
A contextual identity may be based on tasks and relationships.
Browsing activity is a rich source of information for inferring tasks.  For
example, the \texttt{about:profile} Firefox extension uses the Open Directory
Project and Alexa to categorize URLs that the user has
visited~\cite{aboutprofile}.

Tang et al.~show that most people do not take the time to create labels and
maintain friend lists~\cite{tang}. However, the same authors show how to use
heterogeneous networks to infer relationship types (e.g., manager-subordinate)
though the graph itself is unlabeled.

\paragraph{Mitigating implicit linkage}
Given a user's contextual identities, we may be able to prevent information
leakage between them.  For example, preventing third-party cookies from being
read and set across task boundaries means that targeted ads for dating would
not follow the user when visiting a technology news site. Knowing the user's
current contextual identity also allows the browser to suppress URL bar
suggestions that aren't relevant to that identity, protecting the user from
both shoulder-surfing and distraction.
\end{comment}

@@

\paragraph{Expiring posts}
Humans don't have perfect memories, and neither should social
networks~\cite{delete}. Many users already engage in manual auditing and
deletion of old posts~\cite{fbtips2}. A better solution would be to build tools
that let the user manage this more easily through an API: both Twitter and
Facebook provide APIs for post deletion\footnote{Google Plus offers a read-only
API.}.

>>>>

\begin{abstract}
We examine \textit{contextual identity}, a notion that individuals reveal
different aspects of themselves depending on context. At any given time,
an individual may act as a friend, relative, spouse, co-worker, aquaintance, or
stranger. We analyze contextual identity from the perspective of user choice
and control, survey contextual identity violations, and propose new resesarch
directions to enable users to have better control over their contextual
identities.
\end{abstract}

@@

The desire for privacy balances the desire for spontaneous, positive human
interaction, which often requires sharing personal information. The act of
sharing alone does not negate expectation of privacy~\cite{boyd2}.
%The desire for spontaneous, positive human interaction often requires sharing
%personal information. Sharing information can lead to loss of privacy, though
%it does not negate expectation of privacy~\cite{boyd2}. 
Helen Nissenbaum has long argued that privacy violations come not from the
simple sharing of personal information, but rather from sharing that
information in a way that violates social norms~\cite{nissenbaum}.

Lesbian, gay, bisexual, and transgendered youth are at risk for contextual
identity violations.  Access to support is crucial for LGTB-identified youth,
who are at higher risk for bullying and suicide~\cite{hrc}.
Much of that support is only accessible through the Internet in socially
conservative areas. In 2011 Bobbi Duncan and Taylor McCormick, two
students at the University of Texas at Austin, were inadvertantly outed when
their choir director added them to a Facebook group for Queer
Chorus~\cite{fowler}. Even though both students had blocked their parents from
seeing any personal posts revealing their sexuality, because Queer Chorus is a
public group Facebook published notifications to Duncan's and McCormick's
friends, including their parents. Duncan later attempted suicide~\cite{duncan}.

@@

Barth et al.~provide a formal model for contextual
integrity of user data~\cite{barth}. Their work models
the flow of personal information in terms of knowledge states held by agents
acting in given roles, and works well for expressing privacy legislation such as
HIPPA and COPPA. Two limitations of this work are that it does not
model information where the subject involves more than one person, and it
only takes the type of information into account rather than the content or tone
of the information. For example, ``Bob told Alice that Charlie cooked a
delicious dinner'' is out of scope of the model.  In the context of social
networks, information is rarely about a single person, and the tone of the
message is as important to user effect as the contents.

@@

\label{sec:examples}
\subsection{Redistributing information out of context}
Redistributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals and more
to the user's activity stream.  Because of poor opt-out and lack of visibility
into what was being published, user outcry was
immediate~\cite{mccarthy,nakashima}.  In December 2007, Google Reader exposed
RSS feeds of user-marked news stories to the user's Google Talk contacts, which
includes everyone with whom a user has chatted whether they be co-workers,
supervisors, or friends. Although this feed was always public, prior to this
launch it was not discoverable, leading to a bad experience for many
users~\cite{helft}. In February 2010, Google Buzz launched, a product that
exposed the user's most frequent Gmail and Google Talk contacts and their
publicly available (through previously less discoverable) news and photos.  The
combination of exposing contacts (which could include abusive ex-husbands,
co-workers, and friends), as well as aggressively linking photos and news
items, resulted in a huge backlash~\cite{fugitivus,buzz}. In September 2012,
Facebook imported old wall posts from 2008 into the new Timeline interface.
Although wall posts were always visible from profile pages, the new Timeline
interface brought old wall posts, which users used to treat as private messages
before the advent of ``Like'' and comment buttons, to the attention of an
audience that the user never anticipated~\cite{timeline}.

\subsection{Policy errors}
\label{sec:policies}
Some service providers have policies which preclude isolating multiple
contextual identities. For example, Facebook and Google have a ``Real Names''
policy, which requires users to register for accounts with their legal
name.\footnote{Facebook's policy: \url{http://www.facebook.com/help/?page=258984010787183}, Google's policy: \url{http://support.google.com/plus/bin/answer.py?hl=en&answer=1228271}}
%\cite{fb_names,google_names}.  
These policies ignore that
community-building happens in many different contexts, that users have
legitimate reasons for presenting different identities in different contexts,
and that users don't necessarily want those identities to be linked. For
example, disallowing avatar handles as a primary identifier makes building a
gaming community difficult. % Linking a gaming identity to a professional
%identity has already caused problems for 2012 Maine Senate candidate and World
%of Warcraft gamer Colleen Lachowicz~\cite{maine}.
It is impossible to isolate multiple contextual identities in these networks
without violating the terms of service.

@@

Other federated login systems have support for multiple identifiers, but this
feature can be poorly implemented or difficult to discover, leading to low use.
Some login platforms and protocols, such as OAuth and BrowserID, do not suffer
from this policy or design error, giving a user better
control over which contextual identity to assert~\cite{browserid,oauth}.

\subsection{User interface errors}
It is all too easy for users to broadcast information to
an unintended audience. This phenomenon is so common on Twitter that it has its
own name, ``DM fail'', or Direct Message fail: when the user posts a public
message instead of a private, direct message. Representative Anthony Weiner was
a victim of this mistake when he inadvertently published compromising pictures
of himself~\cite{weiner}. Considering that this mistake requires
mistyping a single character (\texttt{@} instead of \texttt{d}), it's no
surprise that DM failures are so common.

@@

\subsection{How do users think about identity?}
We hypothesize many users are not consciously aware of having multiple
contextual identities. In order to develop tools to help users, we need to
understand users' mental models of identity and how information is shared on
the Internet. For computer scientists, authentication is intrinsically linked
to identity, and even in the absence of authentication, using the same device
over time implicitly creates an identity through tracking techniques and local
information. However, these points are far from obvious to a typical user,
especially for users who don't distinguish the web application from the
browser, from the operating system, or from the device.\footnote{What is a
browser? \url{https://www.youtube.com/watch?v=o4MwTvtyrUQ}}

The computer science community must free itself from software-specific
notions of identity if we are to help users to whom the very concept of
identity is an enigma. To understand our users' mental models,
one might ask:

@@

Pew Internet is a good source for phone survey data on privacy and social
media, and danah boyd's work on youth and social media is excellent, but there
is otherwise a dearth of research in these areas~\cite{boyd,pew1,pew2,pew3}.

@@

Using multiple browsers is a useful technique for managing multiple accounts.
Some services support multiple accounts but not multiple login (e.g.,
Twitter). For services that do (e.g., many Google services), the user interface
may be so confusing that it leads to errors, so the best recommendation may be
to use multiple browsers~\cite{yegge}.\footnote{Using multiple
browsers as an alternative to multiple login:
\url{https://support.google.com/accounts/bin/answer.py?hl=en&answer=179235}}
For users who want to separate work and
personal browsing, using multiple browsers is the easiest solution.

@@

Several browsers support separation of profile data, including cookies,
history, passwords, and other local storage. Firefox 
%\url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}.}
and Chrome support multiple profiles, though neither of these is very
discoverable or easy to use.\footnote{Chrome:
\url{https://support.google.com/chrome/bin/answer.py?hl=en&answer=2364824},
Firefox: \url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}} The
threat model does not include users with malicious intent in the same
household.

\paragraph{Private browsing mode}
Private browsing mode exists in all major browsers, but has no standard
behavior~\cite{ABBJ10}.
Interaction with extensions, treatment of cookies, history, and
bookmarks upon entering and exiting are different across browsers.
Similar to multiple profiles, private browsing mode is not secure against local
attacks.

\paragraph{Cookie blockers}
Disconnect and ShareMeNot are browser extensions to disallow interacting with
service providers like Google and Facebook unless users explicitly choose to
interact with that site~\cite{disconnect,franzi}.  Collusion is a tool for
visualizing and blocking cookies~\cite{collusion}, in particular third-party
cookies set by tracking sites which are typically ad networks.

@@

Building reputation online is a long and arduous process; a user who has built
up high credibility in one social network may want to transfer that credibility
when using a new service by using the same identifier or somehow proving in the
new service that they control other credible identities.  The advent
of verified account mechanisms in Twitter, Google Plus and Facebook indicates
that this problem is on the rise.

@@

All five major browsers (Firefox, Chrome, Safari, Internet Explorer, and Opera)
offer synchronizing some set of browser data across devices, with varying
degrees of success. Multiple cross-browser applications also exist to
synchronize bookmarks and passwords.\footnote{Bookmark sync: \url{xmarks.com}, password sync: \url{lastpass.com}}

@@

Given the large number of social network users and that 30-40\% of spoken
communication is devoted to informing others about ourselves, many users are
bound to share information they later regret~\cite{tamir}.

@@

@@

\paragraph{Expiring posts}
Humans don't have perfect memories, and neither should social
networks~\cite{delete}. Many users already engage in manual auditing and
deletion of old posts~\cite{fbtips2}. A better solution would be to build tools
that let the user manage this more easily through an API: both Twitter and
Facebook provide APIs for post deletion.
%\footnote{Google Plus offers a read-only API.}
